UniScatter: Unifying\\Backscatter Communications, Symbiotic Radio and Reconfigurable Intelligent Surface

### 2022-08-21 main.tex
:pencil: Update parameters

	\begin{subsection}{System Model}
		\begin{figure}[!t]
			\centering
			\def\svgwidth{0.8\columnwidth}
			\footnotesize{
				\import{assets/illustration/}{uniscatter_network.eps_tex}
			}
			\caption{A single-user multi-node UniScatter network.}
			\label{fi:uniscatter_network}
		\end{figure}
		To demonstrate the advantages of UniScatter, we consider a single-user multi-node \gls{miso} UniScatter network as shown in Fig. \ref{fi:uniscatter_network}.
		In the primary active point-to-point system, a $Q$-antenna \gls{ap} transmits to a single-antenna user assisted by $K$ nearby single-antenna UniScatter nodes with $M$ reflection states.
		In the secondary backscatter \gls{mac} system, the \gls{ap} serves as the carrier emitter, the distributed UniScatter nodes modulate information over reradiated \gls{rf} signals, and the user decodes their messages.
		For simplicity, we consider a quasi-static block fading model where channels remain constant within coherence block and vary independently between consecutive blocks.
		Due to the physical constraints of load switching speed, we assume the backscatter over primary symbol period ratio is an integer $N \gg 1$.
		We also omit the signal reflected by two or more times and ignore the propagation time difference of different paths.

@@

	\begin{subsection}{Decision Threshold}
		For any given $\{\boldsymbol{p}_k\}_{k \in \mathcal{K}}$ and $\boldsymbol{w}$, problem \eqref{op:weighted_sum_rate} reduces to
		\begin{maxi!}
			{\scriptstyle{\boldsymbol{t}}}{I(x_{\mathcal{K}})}{\label{op:decision_threshold}}{\label{ob:decision_threshold}}
			\addConstraint{\eqref{co:decision_threshold},}
		\end{maxi!}
		which is still non-convex because variable $\boldsymbol{t}$ appears on the limits of integration \eqref{eq:dmtmac}.
		Fortunately, we can further simplify problem \eqref{op:decision_threshold} as a point-to-point rate-optimal quantization design for a discrete-input continuous-output memoryless channel, thanks to Remark \ref{re:backscatter_decision} and \ref{re:augmented_source}.

@@

		Next, we constrain the feasible domain of problem \eqref{op:decision_threshold} from continuous space $\mathbb{R}_+^{L+1}$ to finite candidate set (i.e., fine-grained discrete energy levels) $\mathcal{T}^{L+1}$.
		\begin{figure}[!t]
			\centering
			\resizebox{0.9\columnwidth}{!}{
				\input{assets/illustration/discrete_outputs.tex}
			}
			\caption{
				The decision thresholds are selected from fine-grained discrete energy levels instead of continuous space, and each decision region consists of at least one neighbor energy bins.
			}
			\label{fi:discrete_outputs}
		\end{figure}
		As shown in Fig. \ref{fi:discrete_outputs}, by introducing the extra analog-to-digital conversion step, we can obtain decision regions by grouping adjacent high-resolution energy bins.
		Thus, problem \eqref{op:decision_threshold} is recast as
		\begin{maxi}
			{\scriptstyle{\boldsymbol{t} \in \mathcal{T}^{L+1}}}{I_{\mathrm{B}}(x_{\mathcal{K}};\hat{x}_{\mathcal{K}}),}{\label{op:decision_threshold_recast}}{}
		\end{maxi}
		which can be solved by existing rate-optimal \gls{dmc} quantizer designs.
		For global optimal solutions, \cite{He2021} proposed a \gls{dp} method accelerated by the \gls{smawk} algorithm with computational complexity $\mathcal{O}\bigl(L^2(\mathrm{card}(\mathcal{T})-L)\bigr)$, and \cite{Nguyen2020a} presented a traverse-bisection algorithm using the optimality condition of three neighbor thresholds with complexity $\mathcal{O}\bigl(\mathrm{card}(\mathcal{T})L\log(\mathrm{card}(\mathcal{T})L)\bigr)$.
		In Section \ref{se:simulation_results}, both schemes will be compared with the \gls{ml} thresholding scheme \cite{Qian2019}
		\begin{equation}
			t_{l}^{\mathrm{\gls{ml}}} = N \frac{\sigma_{l-1}^2 \sigma_{l}^2}{\sigma_{l-1}^2 - \sigma_{l}^2} \log \frac{\sigma_{l-1}^2}{\sigma_{l}^2}, \quad l \in \mathcal{L} \setminus \{L\},
			\label{eq:detection_threshold_maximum_likelihood}
		\end{equation}
		which is generally suboptimal for problem \eqref{op:decision_threshold} except when all nodes are with equiprobable inputs.
	\end{subsection}
\end{section}

\begin{section}{Simulation Results}
	% TODO design $\mathcal{T}$
	% TODO add distribution to initializer to smooth curves?
	% TODO compare PGD and primary-MRT beamformers
	In this section, we provide numerical results to evaluate the proposed input, beamforming and decision design over a single-user multi-node UniScatter network.
	We assume the distance between \gls{ap} and user is \qty{10}{\meter}, and $K=2$ UniScatter nodes are uniformly dropped within a disk centered at the user of radius \qty{1}{\meter}.
	% The carrier frequency is $f=\qty{200}{\MHz}$, and we consider \gls{iid} Ricean fading for all channels.
	The carrier frequency is $f=\qty{200}{\MHz}$, and we consider \gls{iid} Ricean fading between all terminals.
	For direct, forward and backward links, we set the path loss exponents to \num{2.6}, \num{2.4} and \num{2}, and the Ricean factor to \num{5}, \num{5} and \num{10}, respectively.
	% For direct, forward and backward links, we choose path loss exponents \num{2.6}, \num{2.4} and \num{2}, and Ricean factor \num{5}, \num{5} and \num{10}, respectively.
	% and the path loss exponents for direct, forward and backward links are \num{2.6}, \num{2.4} and \num{2}, respectively.
	% We consider \gls{iid} Ricean fading for all channels with Ricean factor set to \num{5}, \num{5} and \num{10}
	The \gls{ap} has $Q=4$ antennas with maximum average transmit power $P=\qty{36}{\dBm}$.
	All nodes have amplitude scattering ratio $\alpha=0.5$, symbol period ratio $N=20$, and perform \gls{qam} with $M=4$ input states.
	% and we consider amplitude scattering ratio $\alpha=0.5$ and symbol period ratio $N=20$.
	% The user has a receive antenna gain of \qty{3}{\dBi} and average noise power
	The user is with average noise power $\sigma_v^2=\qty{70}{\dBm}$ and receive antenna gain \qty{3}{\dBi}.
	All rate regions are averaged over at least \num{1000} instances, where ``PSP'' and ``BSP'' means primary and backscatter symbol periods, respectively.
	We choose decision design by \gls{smawk} \cite{He2021} as reference, and select discretize boundaries uniformly over the \qty{95}{\percent} confidence region of edge hypotheses.
	The parameters remain fixed unless otherwise specified.

>>>>

	\begin{subsection}{System Model}
		\begin{figure}[!t]
			\centering
			\def\svgwidth{0.8\columnwidth}
			\footnotesize{
				\import{assets/illustration/}{uniscatter_network.eps_tex}
			}
			\caption{A single-user multi-node UniScatter system.}
			\label{fi:uniscatter_network}
		\end{figure}
		To demonstrate the advantages of UniScatter, we consider a single-user multi-node \gls{miso} UniScatter system as shown in Fig. \ref{fi:uniscatter_network}.
		In the primary active point-to-point system, a $Q$-antenna \gls{ap} transmits to a single-antenna user assisted by $K$ nearby single-antenna UniScatter nodes with $M$ reflection states.
		In the secondary backscatter \gls{mac} system, the \gls{ap} serves as the carrier emitter, the distributed UniScatter nodes modulate information over reradiated \gls{rf} signals, and the user decodes their messages.
		For simplicity, we consider a quasi-static block fading model where channels remain constant within coherence block and vary independently between consecutive blocks.
		Due to the physical constraints of load switching speed, we assume the backscatter over primary symbol period ratio is an integer $N \gg 1$.
		We also omit the signal reflected by two or more times and ignore the propagation time difference of different paths.

@@

	\begin{subsection}{Decision Threshold}
		For any given $\{\boldsymbol{p}_k\}_{k \in \mathcal{K}}$ and $\boldsymbol{w}$, problem \eqref{op:weighted_sum_rate} reduces to
		\begin{maxi!}
			{\scriptstyle{\boldsymbol{t}}}{I(x_{\mathcal{K}})}{\label{op:decision_threshold}}{\label{ob:decision_threshold}}
			\addConstraint{\eqref{co:decision_threshold},}
		\end{maxi!}
		which is still non-convex because variable $\boldsymbol{t}$ appears on the limits of integration \eqref{eq:dmtmac}.
		Fortunately, we can further simplify problem \eqref{op:decision_threshold} as a point-to-point rate-optimal quantizer optimization for a discrete-input continuous-output memoryless channel, thanks to Remark \ref{re:backscatter_decision} and \ref{re:augmented_source}.

@@

		Next, we constrain the feasible domain of problem \eqref{op:decision_threshold} from continuous space $\mathbb{R}_+^{L+1}$ to finite candidate set (i.e., fine-grained discrete energy levels) $\mathcal{T}^{L+1}$.
		\begin{figure}[!t]
			\centering
			\resizebox{0.9\columnwidth}{!}{
				\input{assets/illustration/discrete_outputs.tex}
			}
			\caption{
				The decision thresholds are selected from fine-grained discrete energy levels instead of continuous space, and each decision region consists of at least one neighbor energy bins.
			}
			\label{fi:discrete_outputs}
		\end{figure}
		As shown in Fig. \ref{fi:discrete_outputs}, by introducing the extra analog-to-digital conversion step, we can group adjacent high-resolution energy bins for backscatter decision regions.
		Thus, problem \eqref{op:decision_threshold} is recast as
		\begin{maxi}
			{\scriptstyle{\boldsymbol{t} \in \mathcal{T}^{L+1}}}{I_{\mathrm{B}}(x_{\mathcal{K}};\hat{x}_{\mathcal{K}}),}{\label{op:decision_threshold_recast}}{}
		\end{maxi}
		which can be solved by existing rate-optimal \gls{dmc} quantizer designs.
		For global optimal solutions, \cite{He2021} proposed a \gls{dp} method accelerated by the \gls{smawk} algorithm with computational complexity $\mathcal{O}\bigl(L^2(\mathrm{card}(\mathcal{T})-L)\bigr)$, and \cite{Nguyen2020a} presented a traverse-then-bisect algorithm using the optimality condition of three neighbor thresholds with complexity $\mathcal{O}\bigl(\mathrm{card}(\mathcal{T})L\log(\mathrm{card}(\mathcal{T})L)\bigr)$.
		In Section \ref{se:simulation_results}, both schemes will be compared with the \gls{ml} thresholding scheme \cite{Qian2019}
		\begin{equation}
			t_{l}^{\mathrm{\gls{ml}}} = N \frac{\sigma_{l-1}^2 \sigma_{l}^2}{\sigma_{l-1}^2 - \sigma_{l}^2} \log \frac{\sigma_{l-1}^2}{\sigma_{l}^2}, \quad l \in \mathcal{L} \setminus \{L\},
			\label{eq:detection_threshold_maximum_likelihood}
		\end{equation}
		which is generally suboptimal for problem \eqref{op:decision_threshold} except when all nodes are with equiprobable inputs.
	\end{subsection}
\end{section}

\begin{section}{Simulation Results}
	% TODO compare PGD and primary-MRT beamformers
	In this section, we provide numerical results to evaluate the proposed input distribution, active beamforming, and backscatter decision design for the UniScatter system in Fig. \ref{fi:uniscatter_network}.
	We assume the AP-user distance is \qty{10}{\meter}, $K=2$ UniScatter nodes are randomly dropped within a disk of radius \qty{2}{\meter} centered at the user, the \gls{ap} is with $Q=4$ transmit antennas and an average power budget $P=\qty{36}{\dBm}$, the nodes employs \num{4}-\gls{qam} with amplitude scattering ratio $\alpha=0.5$, the user is with average noise power $\sigma_v^2=\qty{-40}{\dBm}$, and the symbol period ratio is $N=20$.
	For all channels involved, we consider a distance-dependent path loss model
	\begin{equation}
		L(d) = L_0 \biggl(\frac{d_0}{d}\biggr)^\gamma,
	\end{equation}
	together with a Ricean fading model
	\begin{equation}
		\boldsymbol{H} = \sqrt{\frac{\kappa}{1+\kappa}} \bar{\boldsymbol{H}} + \sqrt{\frac{1}{1+\kappa}} \tilde{\boldsymbol{H}}
	\end{equation}
	where $d$ is the transmission distance, $L_0=-\qty{30}{\dB}$ is the reference path loss at distance $d_0=\qty{1}{\meter}$, $\kappa$ is the Ricean $K$-factor, $\bar{\boldsymbol{H}}$ is the deterministic line-of-sight component with unit-magnitude entries, and $\tilde{\boldsymbol{H}}$ is the Rayleigh fading component with entries in standard \gls{iid} \gls{cscg} distribution.
	We choose $\gamma_{\mathrm{D}}=2.6$, $\gamma_{\mathrm{F}}=2.4$, $\gamma_{\mathrm{B}}=2$, and $\kappa_{\mathrm{D}}=\kappa_{\mathrm{F}}=\kappa_{\mathrm{B}}=5$ for direct, forward and backward links.
	The finite threshold domain $\mathcal{T}$ comes from $b$-bit uniform discretization over the critical interval defined by the confidence bounds of edge hypotheses (i.e., lower bound of $\mathcal{H}_1$ and upper bound of $\mathcal{H}_L$) with confidence level $1-\varepsilon$, where $b=8$ and $\varepsilon=\num{2.22e-16}$.
	All achievable rate regions are averaged over \num{3000} realizations, and the parameters remain fixed unless otherwise specified.
